\speculake: Hiding Malware in Speculative Execution

### 2018-04-11 paper/paper.tex
minor paper edits

\newcommand{\speculake}{SpecPiler}

@@

We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
seemingly-benign payload binary. When a separate trigger program performs a specific
pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
executions of the payload binary with similar jumps leads the CPU to mispredict its
branches and speculatively execute the malware's payload, which communicates
results back to the real world via side channels.

@@

\subsection{Branch Prediction}
When a CPU mispredicts a branch, the speculative execution results are
discarded, costing the CPU several cycles as the pipeline is flushed. To
minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
path of execution. These predictors can maintain a short history of previous
branch targets for a particular branch (e.g. whether a certain branch is
frequently taken or not taken, or if it exhibits a particular pattern), and use
this to inform the CPU's guess for speculative execution.

There are two kinds of branches a CPU must be able to predict: \emph{direct} and
\emph{indirect}. A \emph{direct} branch may either jump to a provided address or
continue executing straight through depending on the state of the CPU (e.g. condition
registers). While there are only two potential statically-known targets for a direct
branch, the CPU may not know if the branch is taken or not until preceding
instructions retire. An \emph{indirect} branch is always taken, but its address
is determined by the value of a register or memory address. Direct branches are
typically used for control flow such as \texttt{if} or
\texttt{for}/\texttt{while} statements, while indirect branches are used for
function pointers, class methods, or case statements.

@@

\subsection{Spectre}
- Basic direct branch
- Indirect branch predictor

@@

>>>>

\newcommand{\speculake}{Exspectre}

@@

We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
seemingly-benign payload binary. When a separate trigger program performs a specific
pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
executions of the payload binary with similar jumps causes the CPU to mispredict its
branches and speculatively execute the malware's payload, which communicates
results back to the real world via side channels.

@@

\subsection{Branch Prediction}
When a CPU mispredicts a branch, the speculative execution results are
discarded, costing the CPU several cycles as the pipeline is flushed. To
minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
path of execution. Branch predictors maintain a short history of previous
branch targets for a particular branch (e.g. whether a certain branch is
frequently taken or not taken), and use
this to inform the CPU's guess for speculative execution.

There are two kinds of branches a CPU handles: \emph{direct} and
\emph{indirect}. A \emph{direct} branch may either jump to a provided address or
continue executing straight through depending on the state of the CPU (e.g. condition
registers). While there are only two potential statically-known targets for a direct
branch, the CPU may not know if the branch is taken or not until preceding
instructions retire. An \emph{indirect} branch is always taken, but its address
is determined by the value of a register or memory address. Direct branches are
typically used for control flow such as \texttt{if} or
\texttt{for}/\texttt{while} statements, while indirect branches are used for
function pointers, class methods, or case statements.

@@

\subsection{Spectre}
In early 2018, researchers revealed the Spectre vulnerability, which allows an
attacker to leak information from a victim program.

Spectre uses speculative execution...
- Basic direct branch
- Indirect branch predictor

@@

\speculake malware is comprised of two independent programs: a payload program, and a
trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
exploit, or phishing), a
